Uncertainty propagation for variable (predictor) effects and bias correction in generalized linear (mixed) models???

### 2021-09-28 variable_predictions.Rnw
Some edits on manuscript

Several quantities of interest may be derived from regression models. The first one being the coefficient estimates. Others being predictors, model matrix, predicted values and marginal effects. \emph{predictors} are model input variables and is associated with one or more \emph{variables} (in polynomials, splines and models with interaction). Model matrix refers to the design matrix whose rows include all combination of variables appearing in the interaction terms, along with the ``typical" values of the focal and non-focal predictors.

@@

Figure ~\ref{fig:qoi_pred_plot} shows the comparison of \emph{variable prediction} and unconditional marginal effect of effect of age on household size, based on model estimates in Table ~\ref{tab:qoi_coefs}. Since the model has no interactions, the relationship between the predicted household size and age is linear, hence the marginal effect of age is the slope ($\frac{\Delta \mathrm{hh size}}{\Delta \mathrm{age}}$) of the variable effect line and can be calculated irrespective of the values of wealth index. 
%
\begin{figure}[H]
\begin{center}
<<qoi_pred_plot, echo=FALSE, results=tex, fig=TRUE, message=FALSE>>=
qoi_age_pred_plot
@
\end{center}
\caption{Variable effect plot and marginal effect of age of household head on household size. For a linear model with no interaction, the marginal effect is the slope of the variable effect line.}
\label{fig:qoi_pred_plot}
\end{figure}
%
Variable effect, on the other hand, is the expected household size for a particular age, holding wealth index at its mean. In particular, the purpose and goal of a variable effect seems fairly straightforward; for specified values of (a) focal predictor(s), we want to give a point estimate and confidence intervals for the prediction of the model for a ``typical" (= random sample over the multivariate distribution of non-focal parameters) individual with those values of the predictors.

@@

To get an intuition of how conditioning on the mean values of the non-focal predictors work, suppose we are interested in the variable predictions of a particular predictor (hence forth referred as \emph{focal} predictor otherwise \emph{non-focal}), $x_f$, from the set of predictors. To keep it simple, assume that the model has no interaction terms. Then the idea is to \emph{anchor} the values of non-focal predictors to some particular values. For example fixing the values of \emph{non-focal} predictor(s) at some typical values -- typically determined by averaging (for now) in some meaningful way, for example, arithmetic mean  for continuous and average over the levels of the factors for categorical non-focal predictors. The easiest way to achieve this is by constructing $\bX^\star$ by averaging the columns of non-focal variables in model matrix $\bX$, and together with appropriately chosen values of focal predictor(s).

@@

An alternative, save for later, formulation of variable effect involves, expressing the linear predictor as the sum of the focal and non-focal predictor linear predictor. In particular, $\eta^\star(x_f, \nset{{\bar{x}}}) = \beta_f x_f + \sum \nset{\beta} \nset{{\bar{x}^\star}}$, where $\nset{{\bar{x}^\star}}$ are the appropriately averaged entries of non-focal predictors and $\nset{x}$ is a vector of values of the non-focal predictors for a particular observation.

@@

The computation of $\hat{\eta}^\star$ remains the same as described above. However, to compute $\sigma$, $\Sigma$ is modified by \emph{zeroing-out} (the variance-covariance of all non-focal predictors are assigned zero) variances of non-focal terms. Although this is the simplest approach, it requires centering of continuous predictors prior to model fitting and proper handling of categorical predictors. One way to center predictors is to create a design matrix, $\bX$, and then compute $\bX_{c} = \bX - \bar{\bX}$.

>>>>

Several quantities of interest may be derived from regression models. The first one being the coefficient estimates. Others being \emph{predictors}, \emph{model matrix}, \emph{predicted values} and \emph{marginal effects}. Predictors are model input variables and is associated with one or more \emph{variables} (in polynomials, splines and models with interaction). Model matrix refers to the design matrix whose rows include all combination of variables appearing in the interaction terms, along with the ``typical" values of the focal and non-focal predictors.

@@

Figure ~\ref{fig:qoi_pred_plot} shows the comparison of predictor effect and unconditional marginal effect of age on household size, based on model estimates in Table ~\ref{tab:qoi_coefs}. Since the model has no interactions, the relationship between the predicted household size and age is linear, hence the marginal effect of age is the slope ($\frac{\Delta \mathrm{hh size}}{\Delta \mathrm{age}}$) of the variable effect line and can be calculated irrespective of the values of wealth index. 
%
\begin{figure}[H]
\begin{center}
<<qoi_pred_plot, echo=FALSE, results=tex, fig=TRUE, message=FALSE>>=
qoi_age_pred_plot
@
\end{center}
\caption{Variable effect plot and marginal effect of age of household head on household size. For a linear model with no interaction, the marginal effect is the slope of the variable effect line.}
\label{fig:qoi_pred_plot}
\end{figure}
%
Predictor effect, on the other hand, is the expected household size for a particular age, holding wealth index at its mean. In particular, the purpose and goal of a variable effect seems fairly straightforward; for specified values of (a) focal predictor(s), we want to give a point estimate and confidence intervals for the prediction of the model for a ``typical" (= random sample over the multivariate distribution of non-focal parameters) individual with those values of the predictors.

@@

To get an intuition of how conditioning on the mean values of the non-focal predictors work, suppose we are interested in predictor effects of a particular predictor (hence forth referred as \emph{focal} predictor otherwise \emph{non-focal}), $x_f$, from the set of predictors. To keep it simple, assume that the model has no interaction terms. Then the idea is to \emph{anchor} the values of non-focal predictors to some particular values. For example fixing the values of \emph{non-focal} predictor(s) at some typical values -- typically determined by averaging (for now) in some meaningful way, for example, arithmetic mean  for continuous and average over the levels of the factors for categorical non-focal predictors. The easiest way to achieve this is by constructing $\bX^\star$ by averaging the columns of non-focal variables in model matrix $\bX$, and together with appropriately chosen values of focal predictor(s).

@@

An alternative, save for later, formulation of variable effect involves, expressing the linear predictor as the sum of the focal and non-focal predictor linear predictor. In particular, $\eta^\star(x_f, \nset{{\bar{x}}}) = \beta_f x_f + \sum \nset{\beta} \nset{{\bar{x}^\star}}$, where $\nset{{\bar{x}^\star}}$ are the appropriately averaged entries of non-focal predictors and $x_f$ is a vector of values of the focal predictors for a particular observation.

@@

The computation of $\hat{\eta}^\star$ remains the same as described above. However, to compute $\sigma$, $\Sigma$ is modified by \emph{zeroing-out} (the variance-covariance of all non-focal predictors are assigned zero) variances of non-focal terms. Although this is the simplest approach, it requires centering of continuous predictors prior to model fitting and proper way to average categorical predictors. One way to center predictors is to create a design matrix, $\bX$, and then compute $\bX_{c} = \bX - \bar{\bX}$.
